Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-11-21 paper/paper.tex
added author summary

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
{\color{red} add this}

@@

%%% start IFNrare figure %%%
\begin{figure}[b!]
\centerline{
{\bf \Large A}\includegraphics[width=0.35\textwidth,valign=t]{figures/IFN_stochastic/IFN_reporter/IFN_reporter.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.23\textwidth,valign=t]{figures/IFN_stochastic/Flow/ifn_percent.pdf}
}
\caption{
{\bf Reporter cells to identify and enrich infections that activate IFN expression.}
{\bf (A)} The reporter consists of an IFN promoter that drives expression of a cell-surface protein amenable to MACS and a fluorescent protein.
We created reporters with type I and type III IFN promoters (\SUPPFILE{reporter_sequences}).
In A549 cells, the reporters were efficiently activated by an IFN-inducing strain of Sendai virus (\SUPPFIG{reportervalidation}A).
{\bf (B)}
Frequency of IFN induction upon infection with the influenza virus stock used in the single-cell studies in this paper, as quantified using the type III IFN reporter (see \SUPPFIG{IFNrare_flow} for full flow cytometry data).
The plot also shows uninfected cells, and cells infected with Sendai virus.
The limit of detection of 0.05\% is indicated with a dashed line, and numbers show the median of three measurements.
}
\label{fig:IFNrare}
\end{figure}
%%% end IFNrare figure %%%

@@

%%% start workflow figure
\begin{figure}[b!]
\begin{fullwidth}

@@

\section*{Acknowledgments}
We thank Cole Trapnell, Jason Underwood, Robert Bradley, Daniel Stetson, AJ Velthuis, Katherine Xue, and Hannah Itell for helpful suggestions.
We thank Andy Marty and the Fred Hutch Genomics Core for performing the deep sequencing.
This work was supported by the NIAID of the NIH under grant R01 AI127893 and a Burroughs Wellcome Fund Young Investigator in the Pathogenesis of Infectious Diseases grant to JDB.
ABR was supported by a postdoctoral fellowship from the Damon Runyon Cancer Research Foundation.
JRK was supported by a Washington Research Foundation Undergraduate Research Fellowship and a Mary Gates undergraduate research scholarship from the University of Washington.
JDB is an Investigator of the Howard Hughes Medical Institute.
The funders had no role in study design, data collection and interpretation, or the decision to submit the work for publication.

\section*{Author Contributions}
ABR and JDB designed the project.
ABR performed the experiments with assistance from JRK.
ABR and JDB computationally analyzed the data and wrote the paper.

\section*{Declaration of Interests}
The authors declare that no competing interests exist.

@@

\begin{suppfig}
{\bf \Large A}\includegraphics[width=0.48\textwidth, valign=t]{figures/Validation_Figure/SNP_flow_plot.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.39\textwidth, valign=t]{figures/Validation_Figure/del_flow_plot.pdf}
\caption{
{\bf Flow cytometry data for \FIG{validation}.}
{\bf (A)}
Data for \FIG{validation}A.
A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were infected with stocks of the indicated mutant.
After 13 hours, cells were stained for HA protein and analyzed by FACS.
Contour plots show density of all cells, and IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of uninfected cells in these populations.
For infected cells, the percentage IFN+ was calculated only among the HA+ cells (since these are the ones that are infected).
For the uninfected cells, the percentage IFN+ was calculated among all cells, since uninfected cells do not express HA.
For each viral mutant, two independent stocks were assayed in duplicate (i.e., \#1a and \#1b are replicates of one viral stock, and \#2a and \#2b are replicates of the other).
The infections with replicate \#1 of the wild-type virus were performed at an MOI of 0.1 as determined by TCID50, and all other viruses were infected at an equivalent particle number as determined by HI assay. 
{\bf (B)}
Data for \FIG{validation}B.
The virus with the deletion in PB1 cannot be normalized by HA expression since it expresses lower levels of HA due to the lack of secondary transcription.
Therefore, all cells were infected at an equivalent MOI of 0.3 as determined by TCID50 on MDCK-SIAT1 cells for the wild-type and NS1stop viruses, and on MDCK-SIAT1 cells expressing PB1~\cite{bloom2010permissive} for the PB1del385to2163 virus.
\SUPPFIG{delqPCR} shows that at these equivalent TCID50s, all variants had similar amounts of transcriptionally active virus in the absence of secondary transcription.
The percent IFN+ was calculated for \emph{all} cells (HA+ and HA-) since that is a more fair comparison for the PB1del385to2163 virus.
}
\label{suppfig:validationflow}
\end{suppfig}

>>>>

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Because influenza virus has a high mutation rate, many cells are infected by mutated virions.
But so far, it has been impossible to fully characterize the genetic properties of the virions infecting any given cell, since the conventional techniques of flow cytometry and single-cell RNA-seq only detect if a protein or transcript is present---not whether it has mutations.
Here we develop a new approach that uses long-read PacBio sequencing to determine the full sequences of the virions infecting individual cells.
Most cells are infected by virions that have a mutation or some other defect.
This viral genetic variation explains a substantial amount of the cell-to-cell variability in how much viral mRNA is produced in infected cells, and helps determine whether a cell activates an innate-immune response.
However, our results also show that viral genetic defects do not completely explain cell-to-cell variation in the outcome of infection.
Overall, this study provides the first complete picture of how viral mutations affect the outcome of infection in single cells.

@@

%%% start IFNrare figure %%%
\begin{figure}
\centerline{
{\bf \Large A}\includegraphics[width=0.35\textwidth,valign=t]{figures/IFN_stochastic/IFN_reporter/IFN_reporter.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.23\textwidth,valign=t]{figures/IFN_stochastic/Flow/ifn_percent.pdf}
}
\caption{
{\bf Reporter cells to identify and enrich infections that activate IFN expression.}
{\bf (A)} The reporter consists of an IFN promoter that drives expression of a cell-surface protein amenable to MACS and a fluorescent protein.
We created reporters with type I and type III IFN promoters (\SUPPFILE{reporter_sequences}).
In A549 cells, the reporters were efficiently activated by an IFN-inducing strain of Sendai virus (\SUPPFIG{reportervalidation}A).
{\bf (B)}
Frequency of IFN induction upon infection with the influenza virus stock used in the single-cell studies in this paper, as quantified using the type III IFN reporter (see \SUPPFIG{IFNrare_flow} for full flow cytometry data).
The plot also shows uninfected cells, and cells infected with Sendai virus.
The limit of detection of 0.05\% is indicated with a dashed line, and numbers show the median of three measurements.
}
\label{fig:IFNrare}
\end{figure}
%%% end IFNrare figure %%%

@@

%%% start workflow figure
\begin{figure}
\begin{fullwidth}

@@

\section*{Acknowledgments}
We thank Cole Trapnell, Jason Underwood, Robert Bradley, Daniel Stetson, AJ Velthuis, Katherine Xue, and Hannah Itell for helpful suggestions.
We thank Andy Marty and the Fred Hutch Genomics Core for performing the deep sequencing.
%This work was supported by the NIAID of the NIH under grant R01 AI127893 and a Burroughs Wellcome Fund Young Investigator in the Pathogenesis of Infectious Diseases grant to JDB.
%ABR was supported by a postdoctoral fellowship from the Damon Runyon Cancer Research Foundation.
%JRK was supported by a Washington Research Foundation Undergraduate Research Fellowship and a Mary Gates undergraduate research scholarship from the University of Washington.
%JDB is an Investigator of the Howard Hughes Medical Institute.
%The funders had no role in study design, data collection and interpretation, or the decision to submit the work for publication.

@@

\begin{suppfig}
{\bf \Large A}\includegraphics[width=0.48\textwidth, valign=t]{figures/Validation_Figure/SNP_flow_plot.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.39\textwidth, valign=t]{figures/Validation_Figure/del_flow_plot.pdf}
\caption{
{\bf Flow cytometry data for \FIG{validation}.}
{\bf (A)}
Data for \FIG{validation}A.
A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were infected with stocks of the indicated mutant.
After 13 hours, cells were stained for HA protein and analyzed by FACS.
Contour plots show density of all cells, and IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of uninfected cells in these populations.
For infected cells, the percentage IFN+ was calculated among the HA+ cells (since these are the ones that are infected).
For uninfected cells, the percentage IFN+ was calculated among all cells, since uninfected cells do not express HA.
For each viral mutant, two independent stocks were assayed in duplicate (i.e., \#1a and \#1b are one viral stock, and \#2a and \#2b are the other).
The infections with replicate \#1 of the wild-type virus were performed at an MOI of 0.1 as determined by TCID50, and all other viruses were infected at an equivalent particle number as determined by HI assay. 
{\bf (B)}
Data for \FIG{validation}B.
The virus with the deletion in PB1 cannot be normalized by HA expression since it expresses less HA due to the lack of secondary transcription.
Therefore, all cells were infected at an equivalent MOI of 0.3 as determined by TCID50 on MDCK-SIAT1 cells for wild type and NS1stop, and on MDCK-SIAT1 cells expressing PB1~\cite{bloom2010permissive} for PB1del385to2163.
\SUPPFIG{delqPCR} shows that at these equivalent TCID50s, all variants had similar amounts of transcriptionally active virus in the absence of secondary transcription.
The percent IFN+ was calculated for \emph{all} cells (HA+ and HA-) since that is a more fair comparison for PB1del385to2163.
}
\label{suppfig:validationflow}
\end{suppfig}
